Austrian Parliament: MPs transparency and conflict of interest report

### 2021-02-19 _post/2020-07-21-austrian-parliament-mps-transparency-and-conflict-of-interest-report/austrian-parliament-mps-transparency-and-conflict-of-interest-report.Rmd
without css for perfomanc testing

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693
categories:
  - Austria
  - OCR
  - web scraping

preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false

creative_commons: CC BY-NC-SA

---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T)
```

@@

Recently, at the end of June, the Austrian Parliament published its annual report on MPs' additional sources of revenue and pertaining activities (see [here](https://www.parlament.gv.at/POOL/SWBRETT/ZUSD/BezBegrBVGPar9-NR.pdf){target="_blank"}). The publication is mandatory by law, and basically seeks to curtail potential conflicts of interest and to increase transparency as to MPs' activities outside of parliament. While the list comes, unsurprisingly, with a number of caveats ^[For details see [here.](https://www.parlament.gv.at//POOL/SWBRETT/25020/0010/BezBegrBVGPar9Erkl-NR.pdf#page=1){target="_blank"}], and by itself can hardly constitute a watertight safeguard against the various 'undemocratic temptations', the report nevertheless seems to provide interesting data on MPs' financial statuses and professional machinations. All in all enough reason to give it a closer look, i.e. a blog post.^[Halfway through the blog, I realized that the transparency website meineabgeordneten.at had alreay published a similar, although non-technial blog on this matter. If you speak German, have a look [here](https://www.meineabgeordneten.at/News/detail/Nationalrat_Nebeneinkommen_Offenlegung_2019){target="_blank"}. In a previous version I also had wrongly identified the gender of one MP. Kindly, Marion Breitschopf from meineabgeordneten.at helped clarifying this.] Note that I will focus only on Austria's lower chamber (*Nationalrat*) in this post. If time permits, I'll follow up with one on the upper chamber (*Bundesrat*) at a later point.

@@

# Setup
But step by step:
Let's first take care of the necessary R packages and blog/rmarkdown setup.

<details closed> 
<summary>Code: Load packages</summary> 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
                                      size=3)
library(scales)
library(knitr)
library(paletteer)
library(ggtext)
library(glue)
library(gt)
library(DT)
library(pdftools)
library(lemon)
library(ggiraph)

```
</details>

<details closed> 
<summary>Code: Define rmarkdown options</summary> 
```{r setup, echo = T}
# knit_hooks$set(wrap = function(before, options, envir) {
#   if (before) {
#     paste0("<", options$wrap, ">")
#   } else {
#     paste0("</", options$wrap, ">")
#   }
# })

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "left",
  # fig.height = 5,
  # fig.width = 7,
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  dpi = 300
)

options(width = 180, dplyr.width = 150)
```
</details>

<details closed> 
<summary>Code: Define plot theme</summary> 
```{r echo=TRUE}
theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.margin = margin(l = 0, unit = "cm"),
      plot.title = element_markdown(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),
      axis.text.x = element_text(size = 9, 
                                 color = "grey50"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 9,
                                 color="grey50"),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
}

my_caption=c("data: Austrian Parliament, Liste gemäß § 9 Bezügebegrenzungs-BVG, 2020. \nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")
```
</details>

# Data extraction
As so often, one of the main challenges to actually analyze available data is to 'liberate' the information of interest from the format in which they are provided. The Parliament's report comes as a pretty 'untidy' pdf with subheadings etc.,  a format hardly conducive for further analysis. Here's one MP's declaration from the report:

![](/post/2020-07-03-austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width=100%}

@@

The income categories are:
```{r echo=FALSE}
data.frame(
  category = c(1L, 2L, 3L, 4L, 5L),
  income.bracket = c(
    "1-1,000", "1,001-3,500",
    "3,501-7,000", "7,001-10,000", ">10,000")) %>% 
    gt() %>% 
  gt::cols_label(income.bracket=md("**income bracket (Euro)**"),
                 category=md("**category**")) %>% 
  gt::cols_align(align="right",
                 columns = vars(income.bracket))
```

@@

<details open> 
<summary>Code: Extract text from pdfs </summary> 
```{r}
my_path <- here::here("_blog_data", "MP_activities", "BezBegrBVGPar9-NR.pdf")

df_all <- pdftools::pdf_text(pdf = my_path) %>%
  enframe(.,
    name = NULL,
    value = "text_raw"
  ) %>%
  mutate(text_raw = str_replace_all(text_raw, "\\h+", " "))
```
</details>

The code above provides us with a tibble containing as many rows as the report has pages (186). The expression `str_replace_all(text_raw, "\\h+", " ")` replaces all horizontal white space with a single space, but keeps line breaks (\\n and \\r). The latter will be needed at a later stage when it comes to identifying the location of data of interest and extract them.^[I am grateful to the respondents to my pertaining question on [stackoverflow](https://stackoverflow.com/questions/62678640/stringr-remove-multiple-spaces-but-keep-linebreaks-n-r/62679016#62679016){target="_blank"}]

@@

## Get income category
As a first step, let's extract the name of the MPs and the income category from the raw strings obtained above. The code chunk below is the pertaining code, making use of `stringr` and `regex`. Note the `(?s)` term in the regex expressions. While `.` would match any pattern but not line breaks, the `(?s)` alters this limitation and forces `.` to match also `\n`  and `\r`.^[See [here](https://stackoverflow.com/questions/159118/how-do-i-match-any-character-across-multiple-lines-in-a-regular-expression/45981809#45981809){target="_blank"} for a very helpful answer on SO.]

<details open> 
<summary>Code: Extract MPs' names and income category </summary> 
```{r}
df_all <- df_all %>%
  mutate(name = str_extract(text_raw, regex("(?<=Einkommenskategorie)(?s).*(?=für das Jahr)", ignore_case = T)) %>%
    str_remove(., "\\(i\\)") %>%
    str_trim()) %>%
  mutate(income_category_19_raw = str_extract(text_raw, regex("(?<=2019:\\s?)\\S"))) %>% 
  mutate(income_category_19=case_when(income_category_19_raw=="-" ~ 0,
                                      TRUE ~ as.numeric(income_category_19_raw))) %>% 
    replace_na(., list(income_category_19="missing")) %>%
  select(-income_category_19_raw)

```
</details>

@@

<details open> 
<summary>Code: Account for multi-page declarations </summary>  
```{r}
df_income <- df_all %>%
  select(name, income_category_19) %>%
  distinct()
```

```{r}
nrow(df_all)
nrow(df_income)
```
</details>

## Missing income data
Let's have a first, quick look at the information obtained so far:

<details closed> 
<summary>Code: Number of MPs per income bracket </summary>  
```{r}
tbl_income <- df_income %>% 
  count(income_category_19) %>% 
  gt() %>% 
  gt::cols_label(income_category_19=md("**income category**"),
                 n=md("**number of MPs**")) 
```
</details>

@@

<details closed> 
<summary>Code: MPs with with missing income data</summary>  
```{r}
df_income_missing <- df_income %>% 
  filter(income_category_19=="missing") %>% 
  mutate(family_name=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"))

tbl_missing <- df_income_missing %>% 
  select(name, income_category_19) %>% 
  gt() %>% 
  cols_label(income_category_19=md("**income bracket**"),
             name=md("**MP**")) %>% 
  gt::cols_align(align="center",
                 columns = vars(income_category_19))
```
</details>
```{r echo=FALSE}
tbl_missing
```

@@

<details closed>
<summary>Code: Get entry dates for MPs with missing income data</summary>
```{r eval=FALSE}
library(xml2)
library(rvest)
link <- "https://www.parlament.gv.at/WWER/PARL/J1918/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&NRBR=ALLE&GP=XXVII&R_WF=FR&FR=ALLE&W=W&M=M&listeId=8&FBEZ=FW_008"

@@

```
</details>

@@

<details closed>
<summary>Code: Remove MPs joining outside of reporting period </summary>
```{r echo=TRUE}
nrow(df_income)
df_income <- df_income %>% 
  filter(!name %in% df_income_missing$name)
nrow(df_income)
```
</details>

@@

<details open>
<summary>Code: Get MPs' gender</summary>
```{r}
#get a clean name, without any academic or prefessional titles
df_income <- df_income %>% 
  mutate(name_family=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"),
         .after=name) %>% 
  mutate(name_clean=name %>% str_remove(., regex(",.*$")) %>% 
           str_remove_all(., regex("\\S*\\.")) %>% 
           str_remove_all(., regex("\\(.*\\)")) %>% 
           str_trim())

@@

# manually correct for some missing gender data
df_income<- df_income %>% 
  unnest_wider(name_gender,names_repair="universal") %>% 
  rename(name=name...1) %>% 
  select(-contains("year"), -contains("proportion"), -ends_with("7")) %>% 
  mutate(gender=case_when(str_detect(name, "Brandstötter")==TRUE ~ "female",
                          str_detect(name, "El-Nagashi")==TRUE ~ "female",
                          str_detect(name, "Sibylle")==TRUE ~ "female",
                          str_detect(name, "Himmelbauer")~"female",
                          str_detect(name, "Karlheinz")==TRUE ~ "male",
                          str_detect(name, "Jörg")==TRUE ~ "male",
                          str_detect(name, "Volker")==TRUE ~ "male",
                          str_detect(name, "Gertraud")==TRUE ~ "female",
                          str_detect(name, "Nurten")==TRUE ~ "female",
                          str_detect(name, "Süleyman")==TRUE ~ "male",
                          str_detect(name, "Ribo")==TRUE ~ "female",
                          TRUE ~ as.character(gender)))
```
</details>

## Female share per income group
What the subsequent analysis reveals is that the share of women per income category decreases the further we move up the income ladder. While there is almost gender parity in the income category 0, womens' shares decrease continuously up to category 4.

<details closed>
<summary>Code: Calculate gender ratios and create plots</summary>
```{r}
df_gender <- df_income %>% 
  group_by(income_category_19, gender) %>% 
  summarise(n_obs=n()) %>% 
  mutate(n_rel_category=n_obs/sum(n_obs)) %>% 
  group_by(gender) %>% 
  mutate(n_gender=sum(n_obs)) %>% 
  ungroup() %>% 
  mutate(n_rel_gender=n_obs/n_gender)

@@

```{r}
pl_gender_category <- df_gender %>% 
  # filter(gender=="female") %>% 
  mutate(gender=fct_rev(gender)) %>% 
  ggplot()+
  labs(title=glue("Austrian Nationalrat: Share of <span style  = color:{gender_colors['female']}>female</span> and <span style = color:{gender_colors['male']}>male</span> MPs per income category"),
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of female MPs in brackets. Only lower chamber. Data for 2019.", 105),
       y="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(y=income_category_19,
               x=n_obs,
               fill=gender),
           stat="identity",
           position="fill")+
  geom_text(data=. %>% filter(gender=="female"),
            aes(y=income_category_19,
                x=n_rel_category-0.01,
                label=paste0(round(n_rel_category,2) %>% 
                  percent(scale=100),
                  "\n(",
                  n_obs,
                  ")")),
            color="white",
            lineheight=0.6,
            hjust=1,
            stat="identity")+
  geom_vline(xintercept = .5,
             color="white",
             linetype="dashed"
             )+
  scale_fill_manual(values=gender_colors)+
  scale_x_continuous(label=scales::label_percent(),
                     expand=expansion(mult=c(0, 0.05)))+
  scale_y_discrete(labels=income_categories_short,
                   expand=expansion(mult=c(0, 0)),
                   position = "left")+
  theme_post()+
  theme(legend.position="none",
        plot.title=element_markdown(),
        axis.title.x=element_blank(),
        axis.title.y.left = element_text(angle = 90,
                                         color="grey50",
                                         hjust=1),
        axis.text.y.left = element_text(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        panel.grid.major.y  = element_blank())

```
</details>

```{r echo=FALSE}
pl_gender_category

@@

## Share of income groups per gender group
Another gender sensitive perspective is to look at the distribution of the MPs across income categories within each gender group (in contrast to within each income group as in the previous graph). The graph highlights the difference between female and male MPs in terms of the income categories. While the distribution for female MPs is skewed towards lower income categories, that for male MPs is more concentrated on income category 2 and 3. Furthermore, only 4.4 % of all female MPs have additional incomes falling into the two highest categories, 4 and 5. In contrast, 19 % of all male MPs have income from these categories.

<details closed> 
<summary>Code: graph intra-gender distribution </summary>
```{r}
pl_gender_intra_dist <- df_gender %>% 
  ggplot()+
  labs(title="Austrian Nationalrat: Distribution of MPs over income groups per gender",
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of MPs in brackets. Only lower chamber. Data for 2019.", 105),
       x="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(x=income_category_19,
               y=n_rel_gender,
               fill=gender),
           stat="identity")+
  geom_text(aes(x=income_category_19,
                y=n_rel_gender+0.02,
                label=paste0(percent(n_rel_gender, accuracy = 0.1),
                             "\n",
                             "(", n_obs, ")")),
            lineheight=0.7,
            stat="identity")+
  scale_fill_manual(values=gender_colors, 
                    na.value = "grey30")+
  scale_x_discrete(labels=income_categories_short)+
  scale_y_continuous(label=scales::label_percent(accuracy=1),
                     expand=expansion(mult=c(0, 0.1)))+
  facet_wrap(vars(gender))+
  theme_post()+
  theme(legend.position = "none",
        axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),)
```
</details>

@@

## Get MPs' party membership
As the next step to facilitate the analysis of the data, let's add MPs' party affiliation. I do this by importing the pertaining data from wikipedia's entry on Austria's current MP (the parliament's website was simply to messy for my liking). Here, the `rvest` package is our friend. Once the data on MP membership is retrieved, I merge it with MPs' income/transparency data. The key/identifier to merge both sets of data is an MP's name. To make this work, I had first to remove all academic/professional titles (it's Austria after all) and harmonize the sequence of first and family names.

<details closed> 
<summary>Code: Get party affiliation </summary>  
```{r}
wiki_link <- "https://de.wikipedia.org/wiki/Liste_der_Abgeordneten_zum_%C3%96sterreichischen_Nationalrat_(XXVII._Gesetzgebungsperiode)"

@@

<details closed> 
<summary>Code: plots on income categories</summary>  
```{r}
df_agg <- df_income %>% 
  ungroup() %>%
  mutate(income_category_19_fac=as_factor(income_category_19) %>% 
           fct_inseq()) %>% 
  mutate(party=fct_infreq(party)) %>% 
  group_by(income_category_19, party, .drop = F) %>% 
  summarise(n_obs=n()) %>% 
  group_by(party) %>% 
  mutate(n_obs_total=sum(n_obs)) %>% 
  mutate(n_obs_rel=n_obs/n_obs_total)

>>>>

author:
  - first_name: "Roland"
    last_name: "Schmidt"
    url: https://werk.statt.codes
    orcid_id: 0000-0002-6897-1693

categories:
  - Austria
  - OCR
  - web scraping

# preview: austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/preview.png

output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
    code_folding: TRUE

creative_commons: CC BY-NC-SA
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = T,
                      dev = "png")
```

@@

Recently, at the end of June, the Austrian Parliament published its annual report on MPs' additional sources of revenue and pertaining activities (see [here](https://www.parlament.gv.at/POOL/SWBRETT/ZUSD/BezBegrBVGPar9-NR.pdf){target="_blank"}). The publication is mandatory by law, and basically seeks to curtail potential conflicts of interest and to increase transparency as to MPs' activities outside of parliament. While the list comes, unsurprisingly, with a number of caveats [^1], and by itself can hardly constitute a watertight safeguard against the various 'undemocratic temptations', the report nevertheless seems to provide interesting data on MPs' financial statuses and professional machinations. All in all enough reason to give it a closer look, i.e. a blog post.[^2] Note that I will focus only on Austria's lower chamber (*Nationalrat*) in this post. If time permits, I'll follow up with one on the upper chamber (*Bundesrat*) at a later point.

[^1]: For details see [here.](https://www.parlament.gv.at//POOL/SWBRETT/25020/0010/BezBegrBVGPar9Erkl-NR.pdf#page=1){target="_blank"}

[^2]: Halfway through the blog, I realized that the transparency website meineabgeordneten.at had alreay published a similar, although non-technial blog on this matter. If you speak German, have a look [here](https://www.meineabgeordneten.at/News/detail/Nationalrat_Nebeneinkommen_Offenlegung_2019){target="_blank"}. In a previous version I also had wrongly identified the gender of one MP. Kindly, Marion Breitschopf from meineabgeordneten.at helped clarifying this.

@@

# Setup

But step by step: Let's first take care of the necessary R packages and blog/rmarkdown setup.

<details closed>

<summary>Code: Load packages</summary>

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(extrafont)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
# hrbrthemes::update_geom_font_defaults(family = "Roboto Condensed",
                                      # size=3)
library(scales)
library(knitr)
library(paletteer)
library(ggtext)
library(glue)
library(gt)
library(DT)
library(pdftools)
library(lemon)
library(ggiraph)

</details>

<details closed>

<summary>Code: Define rmarkdown options</summary>

```{r setup, echo = T}
# knit_hooks$set(wrap = function(before, options, envir) {
#   if (before) {
#     paste0("<", options$wrap, ">")
#   } else {
#     paste0("</", options$wrap, ">")
#   }
# })

# knitr::opts_chunk$set(
#   echo = TRUE,
#   fig.align = "left",
#   # fig.height = 5,
#   # fig.width = 7,
#   out.width = "100%",
#   message = FALSE,
#   warning = FALSE
# )
# 
# options(width = 180, dplyr.width = 150)
```

</details>

<details closed>

<summary>Code: Define plot theme</summary>

```{r echo=F}
plot_bg_color <- readr::read_file(file=here::here("theme.css")) %>% 
  str_extract(., regex("(?<=blog-bg-color:).*?(?=;)")) %>%
  str_trim() %>% 
  str_extract(., regex("^#\\S+"))

# theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() +
    theme(
      plot.background = element_rect(fill = plot_bg_color, color=NA),
      panel.background = element_rect(fill = plot_bg_color, color=NA),
      plot.margin = margin(l = 0, unit = "cm"),
      plot.title = element_text(
        color = "grey20",
        face = "bold",
        margin = margin(l = 0, unit = "cm"),
        size = 13
      ),
      plot.title.position = "plot",
      plot.subtitle = element_text(
        color = "grey50",
        margin = margin(t = 0.2, b = 0.3, unit = "cm"),
        size = 11
      ),
      plot.caption = element_text(
        color = "grey50",
        size = 8,
        hjust = c(0)
      ),
      plot.caption.position = "panel",
      axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),
      axis.text.x = element_text(size = 9, 
                                 color = "grey50"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 9,
                                 color="grey50"),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0.25, "cm"),
      panel.spacing.y = unit(0.25, "cm"),
      strip.text = element_text(
        angle = 0,
        size = 9,
        vjust = 1,
        face = "bold"
      ),
      legend.title = element_text(
        color = "grey30",
        face = "bold",
        vjust = 1,
        size = 7
      ),
      legend.text = element_text(
        size = 7,
        color = "grey30"
      ),
      legend.justification = "left",
      legend.box = "horizontal", # arrangement of multiple legends
      legend.direction = "vertical",
      legend.margin = margin(l = 0, t = 0, unit = "cm"),
      legend.spacing.y = unit(0.07, units = "cm"),
      legend.text.align = 0,
      legend.box.just = "top",
      legend.key.height = unit(0.2, "line"),
      legend.key.width = unit(0.5, "line"),
      text = element_text(size = 5)
    )
#}

theme_post <- function() {
  hrbrthemes::theme_ipsum_rc() 
}

my_caption=c("data: Austrian Parliament, Liste gemäß § 9 Bezügebegrenzungs-BVG, 2020. \nanalysis: Roland Schmidt | @zoowalk | https://werk.statt.codes")
```

</details>

# Data extraction

As so often, one of the main challenges to actually analyze available data is to 'liberate' the information of interest from the format in which they are provided. The Parliament's report comes as a pretty 'untidy' pdf with subheadings etc., a format hardly conducive for further analysis. Here's one MP's declaration from the report:

![](austrian-parliament-mps-transparency-and-conflict-of-interest-report_files/list_example - Copy.PNG){width="100%"}

@@

The income categories are:

```{r echo=FALSE}
data.frame(
  category = c(1L, 2L, 3L, 4L, 5L),
  income.bracket = c(
    "1-1,000", "1,001-3,500",
    "3,501-7,000", "7,001-10,000", ">10,000")) %>% 
    gt() %>% 
  gt::cols_label(income.bracket=md("**income bracket (Euro)**"),
                 category=md("**category**")) %>% 
  gt::cols_align(align="right",
                 columns = vars(income.bracket))
```

@@

<details open>

<summary>Code: Extract text from pdfs </summary>

```{r}
my_path <- here::here("_blog_data", "MP_activities", "BezBegrBVGPar9-NR.pdf")

df_all <- pdftools::pdf_text(pdf = my_path) %>%
  enframe(.,
    name = NULL,
    value = "text_raw"
  ) %>%
  mutate(text_raw = str_replace_all(text_raw, "\\h+", " "))
```

</details>

The code above provides us with a tibble containing as many rows as the report has pages (186). The expression `str_replace_all(text_raw, "\\h+", " ")` replaces all horizontal white space with a single space, but keeps line breaks (\\n and \\r). The latter will be needed at a later stage when it comes to identifying the location of data of interest and extract them.[^3]

[^3]: I am grateful to the respondents to my pertaining question on [stackoverflow](https://stackoverflow.com/questions/62678640/stringr-remove-multiple-spaces-but-keep-linebreaks-n-r/62679016#62679016){target="_blank"}

@@

## Get income category

As a first step, let's extract the name of the MPs and the income category from the raw strings obtained above. The code chunk below is the pertaining code, making use of `stringr` and `regex`. Note the `(?s)` term in the regex expressions. While `.` would match any pattern but not line breaks, the `(?s)` alters this limitation and forces `.` to match also `\n` and `\r`.[^4]

[^4]: See [here](https://stackoverflow.com/questions/159118/how-do-i-match-any-character-across-multiple-lines-in-a-regular-expression/45981809#45981809){target="_blank"} for a very helpful answer on SO.

<details open>

<summary>Code: Extract MPs' names and income category </summary>

```{r}
df_all <- df_all %>%
  mutate(name = str_extract(text_raw, regex("(?<=Einkommenskategorie)(?s).*(?=für das Jahr)", ignore_case = T)) %>%
    str_remove(., "\\(i\\)") %>%
    str_trim()) %>%
  mutate(income_category_19_raw = str_extract(text_raw, regex("(?<=2019:\\s?)\\S"))) %>% 
  mutate(income_category_19=case_when(income_category_19_raw=="-" ~ 0,
                                      TRUE ~ as.numeric(income_category_19_raw))) %>% 
    replace_na(., list(income_category_19="missing")) %>%
  select(-income_category_19_raw)

</details>

@@

<details open>

<summary>Code: Account for multi-page declarations </summary>

```{r}
df_income <- df_all %>%
  select(name, income_category_19) %>%
  distinct()
```

```{r}
nrow(df_all)
nrow(df_income)
```

</details>

## Missing income data

Let's have a first, quick look at the information obtained so far:

<details closed>

<summary>Code: Number of MPs per income bracket </summary>

```{r}
tbl_income <- df_income %>% 
  count(income_category_19) %>% 
  gt() %>% 
  gt::cols_label(income_category_19=md("**income category**"),
                 n=md("**number of MPs**")) 
```

</details>

@@

<details closed>

<summary>Code: MPs with with missing income data</summary>

```{r}
df_income_missing <- df_income %>% 
  filter(income_category_19=="missing") %>% 
  mutate(family_name=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"))

tbl_missing <- df_income_missing %>% 
  select(name, income_category_19) %>% 
  gt() %>% 
  cols_label(income_category_19=md("**income bracket**"),
             name=md("**MP**")) %>% 
  gt::cols_align(align="center",
                 columns = vars(income_category_19))
```

</details>

```{r echo=FALSE}
tbl_missing
```

@@

<details closed>

<summary>Code: Get entry dates for MPs with missing income data</summary>

```{r eval=FALSE}
library(xml2)
library(rvest)
link <- "https://www.parlament.gv.at/WWER/PARL/J1918/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&NRBR=ALLE&GP=XXVII&R_WF=FR&FR=ALLE&W=W&M=M&listeId=8&FBEZ=FW_008"

@@

</details>

@@

<details closed>

<summary>Code: Remove MPs joining outside of reporting period </summary>

```{r echo=TRUE}
nrow(df_income)
df_income <- df_income %>% 
  filter(!name %in% df_income_missing$name)
nrow(df_income)
```

</details>

@@

<details open>

<summary>Code: Get MPs' gender</summary>

```{r}
#get a clean name, without any academic or prefessional titles
df_income <- df_income %>% 
  mutate(name_family=str_remove(name, ",.*$") %>% 
           str_extract(., "\\S*$"),
         .after=name) %>% 
  mutate(name_clean=name %>% str_remove(., regex(",.*$")) %>% 
           str_remove_all(., regex("\\S*\\.")) %>% 
           str_remove_all(., regex("\\(.*\\)")) %>% 
           str_trim())

@@

# manually correct for some missing gender data
df_income<- df_income %>% 
  unnest_wider(name_gender,names_repair="universal") %>% 
  rename(name=name...1) %>% 
  select(-contains("year"), -contains("proportion"), -ends_with("7")) %>% 
  mutate(gender=case_when(str_detect(name, "Brandstötter")==TRUE ~ "female",
                          str_detect(name, "El-Nagashi")==TRUE ~ "female",
                          str_detect(name, "Sibylle")==TRUE ~ "female",
                          str_detect(name, "Himmelbauer")~"female",
                          str_detect(name, "Karlheinz")==TRUE ~ "male",
                          str_detect(name, "Jörg")==TRUE ~ "male",
                          str_detect(name, "Volker")==TRUE ~ "male",
                          str_detect(name, "Gertraud")==TRUE ~ "female",
                          str_detect(name, "Nurten")==TRUE ~ "female",
                          str_detect(name, "Süleyman")==TRUE ~ "male",
                          str_detect(name, "Ribo")==TRUE ~ "female",
                          TRUE ~ as.character(gender)))
```

</details>

## Female share per income group

What the subsequent analysis reveals is that the share of women per income category decreases the further we move up the income ladder. While there is almost gender parity in the income category 0, womens' shares decrease continuously up to category 4.

<details closed>

<summary>Code: Calculate gender ratios and create plots</summary>

```{r}
df_gender <- df_income %>% 
  group_by(income_category_19, gender) %>% 
  summarise(n_obs=n()) %>% 
  mutate(n_rel_category=n_obs/sum(n_obs)) %>% 
  group_by(gender) %>% 
  mutate(n_gender=sum(n_obs)) %>% 
  ungroup() %>% 
  mutate(n_rel_gender=n_obs/n_gender)

@@

```{r}
pl_gender_category <- df_gender %>% 
  # filter(gender=="female") %>% 
  mutate(gender=fct_rev(gender)) %>% 
  ggplot()+
  labs(title=glue("Austrian Nationalrat: Share of <span style  = color:{gender_colors['female']}>female</span> and <span style = color:{gender_colors['male']}>male</span> MPs per income category"),
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of female MPs in brackets. Only lower chamber. Data for 2019.", 105),
       y="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(y=income_category_19,
               x=n_obs,
               fill=gender),
           stat="identity",
           position="fill")+
  geom_text(data=. %>% filter(gender=="female"),
            aes(y=income_category_19,
                x=n_rel_category-0.01,
                label=paste0(round(n_rel_category,2) %>% 
                  percent(scale=100),
                  "\n(",
                  n_obs,
                  ")")),
            color="white",
            lineheight=0.6,
            hjust=1,
            stat="identity")+
  geom_vline(xintercept = .5,
             color="white",
             linetype="dashed"
             )+
  scale_fill_manual(values=gender_colors)+
  scale_x_continuous(label=scales::label_percent(),
                     expand=expansion(mult=c(0, 0.05)))+
  scale_y_discrete(labels=income_categories_short,
                   expand=expansion(mult=c(0, 0)),
                   position = "left")+
  theme_post()#+
  # theme(legend.position="none",
  #       plot.title=element_markdown(),
  #       axis.title.x=element_blank(),
  #       axis.title.y.left = element_text(angle = 90,
  #                                        color="grey50",
  #                                        hjust=1),
  #       axis.text.y.left = element_text(),
  #       axis.text.y.right = element_blank(),
  #       axis.title.y.right = element_blank(),
  #       panel.grid.major.y  = element_blank())

pl_gender_category

@@

</details>

```{r echo=FALSE, fig.height=6, fig.width=10, include=T, device="ragg_png"}
pl_gender_category

## Share of income groups per gender group

Another gender sensitive perspective is to look at the distribution of the MPs across income categories within each gender group (in contrast to within each income group as in the previous graph). The graph highlights the difference between female and male MPs in terms of the income categories. While the distribution for female MPs is skewed towards lower income categories, that for male MPs is more concentrated on income category 2 and 3. Furthermore, only 4.4 % of all female MPs have additional incomes falling into the two highest categories, 4 and 5. In contrast, 19 % of all male MPs have income from these categories.

<details closed>

<summary>Code: graph intra-gender distribution </summary>

```{r}
pl_gender_intra_dist <- df_gender %>% 
  ggplot()+
  labs(title="Austrian Nationalrat: Distribution of MPs over income groups per gender",
       subtitle=str_wrap("Income: average monthly income from professional positions in addition to MPs' salary. Absolute number of MPs in brackets. Only lower chamber. Data for 2019.", 105),
       x="[income category] average monthly income",
       caption=my_caption)+
  geom_bar(aes(x=income_category_19,
               y=n_rel_gender,
               fill=gender),
           stat="identity")+
  geom_text(aes(x=income_category_19,
                y=n_rel_gender+0.02,
                label=paste0(percent(n_rel_gender, accuracy = 0.1),
                             "\n",
                             "(", n_obs, ")")),
            lineheight=0.7,
            stat="identity")+
  scale_fill_manual(values=gender_colors, 
                    na.value = "grey30")+
  scale_x_discrete(labels=income_categories_short)+
  scale_y_continuous(label=scales::label_percent(accuracy=1),
                     expand=expansion(mult=c(0, 0.1)))+
  facet_wrap(vars(gender))+
  # theme_post()+
  theme(legend.position = "none",
        axis.title.x = element_text(angle = 0,
                                         color="grey50",
                                         hjust=1),)
```

</details>

@@

```{r}
mtcars %>% 
  ggplot()+
  geom_point(aes(
    x=mpg,
    y=disp
  ))
```

## Get MPs' party membership

As the next step to facilitate the analysis of the data, let's add MPs' party affiliation. I do this by importing the pertaining data from wikipedia's entry on Austria's current MP (the parliament's website was simply to messy for my liking). Here, the `rvest` package is our friend. Once the data on MP membership is retrieved, I merge it with MPs' income/transparency data. The key/identifier to merge both sets of data is an MP's name. To make this work, I had first to remove all academic/professional titles (it's Austria after all) and harmonize the sequence of first and family names.

<details closed>

<summary>Code: Get party affiliation </summary>

```{r}
wiki_link <- "https://de.wikipedia.org/wiki/Liste_der_Abgeordneten_zum_%C3%96sterreichischen_Nationalrat_(XXVII._Gesetzgebungsperiode)"

@@

<details closed>

<summary>Code: plots on income categories</summary>

```{r}
df_agg <- df_income %>% 
  ungroup() %>%
  mutate(income_category_19_fac=as_factor(income_category_19) %>% 
           fct_inseq()) %>% 
  mutate(party=fct_infreq(party)) %>% 
  group_by(income_category_19, party, .drop = F) %>% 
  summarise(n_obs=n()) %>% 
  group_by(party) %>% 
  mutate(n_obs_total=sum(n_obs)) %>% 
  mutate(n_obs_rel=n_obs/n_obs_total)
